---
title: "Untitled"
author: "Blake Shurtz"
date: "February 14, 2019"
output: pdf_document
---
####First inning (prior)

```{r, eval=FALSE}
###read in pre-game data
library(readr)
wspre <- read_csv("data/WS Investigation/wsg5prepbp.csv")
wspre <- wsg5prepbp
###if NA's, try below
wspre <- as.data.frame(wspre)
wspre$home <- as.factor(wspre$home)
wspre$away <- as.factor(wspre$away)
```

```{r}
TEMP <- wspre[1,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```
LAD has higher chance. (Note, the models need to be updated)

####First Inning, 1st half

```{r}
TEMP <- wspre[2,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```


####First Inning, 2nd half

```{r}
TEMP <- wspre[3,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

OK, weird. Probability of home team winning increases despite Boston having a better inning.

####2nd inning (prior)

```{r}
TEMP <- wspre[4,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####2nd inning (2nd half)

```{r}
TEMP <- wspre[5,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

Weird, why?

####3rd inning (prior)

```{r}
TEMP <- wspre[6,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####3rd inning (2nd half)

```{r}
TEMP <- wspre[7,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

Big drop is batting average for BOS while LAD maintains. Basically, LAD is getting more hits.

####4th inning (prior)

```{r}
TEMP <- wspre[8,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####4th inning (2nd half)

```{r}
TEMP <- wspre[9,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####5th inning (prior)

```{r}
TEMP <- wspre[10,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####5th inning (2nd half)

```{r}
TEMP <- wspre[11,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####6th inning (prior)

```{r}
TEMP <- wspre[12,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####6th inning (2nd half)

```{r}
TEMP <- wspre[13,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####7th inning (prior)

```{r}
TEMP <- wspre[14,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####7th inning (2nd half)

```{r}
TEMP <- wspre[15,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####8th inning (prior)

```{r}
TEMP <- wspre[16,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####8th inning (2nd half)

```{r}
TEMP <- wspre[17,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

####9th inning (prior)

```{r}
TEMP <- wspre[18,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```

Drop is BOS batting average, model doesn't assume ending at 9 innings.

####9th inning (2nd half)

```{r}
TEMP <- wspre[19,] #rdiff is how they performed as the home team
head(TEMP)
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model); prob_success
```