---
title: "Bayesian Baseball"
author: "Blake Shurtz"
date: "February 5, 2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
library(RMariaDB)
con <- dbConnect(MariaDB(),
                 user = 'guest',
                 password = 'password',
                 host = 'mydbinstance4.c1uducbod6js.us-west-1.rds.amazonaws.com',
                 dbname='bayesianbaseball')
mydata2 <- dbReadTable(conn = con, name = 'mydata', value = mydata2, overwrite = FALSE)

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
library(tidyverse)
library(reshape)
mydata <- read_csv("data/mydata.csv")
mydata <- as.data.frame(mydata)
namesmydata <- colnames(mydata)
attributes(mydata) <- NULL
mydata <- as.data.frame(mydata)
colnames(mydata) <- namesmydata
colnames(mydata)[1] <- "X"
```

OPEN modelexecution2.RDATA.

###The Model  

```{r, message=FALSE, eval=FALSE}
set.seed(1234)
options(mc.cores = parallel::detectCores())
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')

vivs.m <- map2stan(
  alist(
    rdiff ~ dnorm( mu, sigma),
    ###likelihood function, a function of multiple linear models
    mu <- A + BH + PH + BA + PA + FH + FA,
    #varying intercepts
    A <- h[home] + a[away] + a, 
    #home- batting
    BH <- hits_h[home] * Hh + double_h[home] * B2Bh + triple_h[home] * B3Bh + 
         HR_h[home] * BHRh + balls_h[home] * BBBh + bh, 
    #home- pitching
    PH <-  hits_allowed_h[home] * HHAh + pballs_h[home] * PBBh + 
          pstrikeouts_h[home] * PSOh +  strikes_h[home] * Strh + ph,
    #away- batting
    BA <- hits_a[away] * Ha + double_a[away] * B2Ba + triple_a[away] * B3Ba + 
      HR_a[away] * BHRa + balls_a[away] * BBBa +  ba, 
    #away- pitching
    PA <- hits_allowed_a[away] * HHAa + pballs_a[away] * PBBa + 
      pstrikeouts_a[away] * PSOa + strikes_a[away] * Stra + pa,
    #home & away fielding- group-level parameters
    FH <- fh1 * PO_h + fh2 * A_h + fh3 * E_h + fh4 * DP_h + fh,
    FA <-  fa1 * PO_a + fa2 * A_a + fa3 * E_a + fa4 * DP_a + fa,
    ###adaptive priors for the correlation matrix
    c(h, hits_h, double_h, triple_h, HR_h, balls_h, hits_allowed_h, pballs_h, 
      pstrikeouts_h, strikes_h)[home] ~ dmvnormNC(sigma_home,Rho_home),
    c(a, hits_a, double_a, triple_a, HR_a, balls_a, hits_allowed_a, pballs_a, 
      pstrikeouts_a, strikes_a)[away] ~ dmvnormNC(sigma_away,Rho_away),
    ###standard priors
    c(a, bh, ph, fh, ba, pa, fa) ~ dnorm(0,1), #hyperpriors
    c(fh1, fh2, fh3, fh4, fa1, fa2, fa3, fa4) ~ dnorm(0,1), #scaled variables
    sigma_home ~ dcauchy(0,2),
    sigma_away ~ dcauchy(0,2),
    Rho_home ~ dlkjcorr(4),
    Rho_away ~ dlkjcorr(4),
    sigma ~ dcauchy(0,2)),
  data=mydata,
  iter=10000, warmup=1000, chains=1, cores=4, control = list(adapt_delta = 0.99, max_treedepth = 15))

```

###Diagnostics

```{r traceplot, fig.width=10, fig.height=4}
post <- extract.samples(vivs.m, inc_warmup=TRUE)
TEMP <- post$sigma
TEMP <- as.data.frame(TEMP)
TEMP$index <- seq(1:nrow(TEMP))
ggplot(TEMP, aes(x=index, y=TEMP)) + geom_line() + theme_minimal() +
  ylab("sigma") + xlab("") + ggtitle("Traceplot of Sigma") +
  geom_rect(aes(xmin=0, xmax=200, ymin=2, ymax=2.25), alpha=0.005)
```

```{r neff and rhat, warning=FALSE}
library(latex2exp)
###
TEMP <- precis(vivs.m, depth=3)
###
ggplot(TEMP, aes(n_eff)) + geom_histogram(binwidth=250, fill="white", color="black") + theme_minimal() +
  xlab("Number of Effective Samples") + ylab("Frequency") + ggtitle("Efficiency of Chain Convergence")
###
ggplot(TEMP, aes(Rhat)) + geom_histogram(binwidth=.00025, fill="white", color="black") + theme_minimal() +
  xlab(TeX("$\\hat{R}$")) + ylab("Frequency") + ggtitle("Gelman-Rubin Convergence Diagnostic")

```

###Data, Priors, Posteriors
```{r}
ggplot(mydata, aes(rdiff)) + geom_histogram(binwidth=1, fill="white", color="black") + theme_minimal() +
  xlab(("Difference in Scores (Home - Away)")) + ylab("Frequency") + ggtitle("Distribution of Difference in Runs")
```


```{r priorpostsigma, message=FALSE, echo=FALSE, warning=FALSE}
PRIOR <- extract.prior(vivs.m, n=750)
POST <- extract.samples(vivs.m, n=750)

#prior for sigma instead
sigma <- PRIOR$sigma; sigma <- as.data.frame(sigma); sigma$class <- rep("prior", nrow(sigma))
#post for sigma
sigma.po <- POST$sigma; sigma.po <- as.data.frame(sigma.po); sigma.po$class <- rep("post", nrow(sigma.po))
colnames(sigma.po) <- c("sigma", "class")
TEMP <- as.data.frame(rbind(sigma, sigma.po))

ggplot(TEMP) + 
  geom_density(aes(x = sigma, linetype=class), kernel="gaussian") +
  xlim(-1, 5) +
  theme_minimal() +
  theme(legend.position="bottom") +
  ggtitle("Prior and Posterior For Sigma")
```

###Correlation Matrix
```{r scatterplotrho, echo=FALSE, fig.width=6, fig.height=6, warning=FALSE}
###HOME
R <- rlkjcorr(750, K=3, eta=4) #750, consistent with post samples
R <- as.data.frame(R)
TEMP <- NULL
#prior
TEMP$home <- as.data.frame(R[,2,2]); #random correlations with eta=4
TEMP$away <- as.data.frame(R[,3,2]); #random correlations with eta=4
TEMP <- as.data.frame(TEMP)
colnames(TEMP) <- c("home", "away")
#post
post <- extract.samples(vivs.m)
TEMP2 <- NULL
TEMP2$home <- melt(post$Rho_home[,1,2]); ###correlation between intercepts and slopes (hits, home)
TEMP2$away <- melt(post$Rho_away[,1,2]); ###same as above, but away (still hits)
TEMP2 <- as.data.frame(TEMP2)
colnames(TEMP2) <- c("home", "away")
#bind
TEMP <- rbind(TEMP, TEMP2)
#group
TEMP$group <- c(rep("prior", nrow(TEMP)/2), rep("posterior", nrow(TEMP)/2))
#describe
#head(TEMP)
#dim(TEMP)
#plot
ggplot(TEMP, (aes(home, away))) +
  geom_point(aes(alpha=rev(group)), size=.75) +
  theme_minimal() +
  theme(legend.position="bottom") +
  ggtitle("Correlation between Intercepts and Slopes")
```

```{r scatterplot alternative}
TEMP$B <- rev(TEMP$group)
ggplot(TEMP, (aes(home, away))) +
  geom_vline(xintercept=0, col="gray", size=.2) +
  geom_hline(yintercept=0, col="gray", size=.2) +
  geom_density2d(aes(alpha=B), col = "black", size=.2) +
  theme_minimal() +
  theme(legend.position="bottom") +
  ggtitle("Correlation between Intercepts and Slopes") +
  xlab("Home Team") + ylab("Away Team") 
```


###Model Comparison
```{r, echo=FALSE}
TEMP <- mydata
TEMP$home <- coerce_index(TEMP$home)
TEMP$away <- coerce_index(TEMP$away)
```

```{r nonmlm.m,eval=FALSE, echo=FALSE}
set.seed(1234)
options(mc.cores = parallel::detectCores())
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')

nonmlm.m <- map2stan(
  alist(
    rdiff ~ dnorm( mu, sigma),
    mu <- inter +
      h * home + a * away + 
      hits_h * Hh + double_h * B2Bh + triple_h * B3Bh + HR_h * BHRh + balls_h * BBBh +
      hits_a * Ha + double_a * B2Ba + triple_a * B3Ba + HR_a * BHRa + balls_a * BBBa +
      hits_allowed_h * HHAh + pballs_h * PBBh + pstrikeouts_h * PSOh +  strikes_h * Strh +
      hits_allowed_a * HHAa + pballs_a * PBBa + pstrikeouts_a * PSOa + strikes_a * Stra +
      fh1 * PO_h + fh2 * A_h + fh3 * E_h + fh4 * DP_h + fa1 * PO_a + fa2 * A_a + fa3 * E_a + fa4 * DP_a,
    ###adaptive priors for coefficients
    inter ~ dunif(-100, 100),
    h ~ dunif(-100, 100),
    a ~ dunif(-100, 100),
    ###
    hits_h ~ dunif(-100, 100),
    hits_a ~ dunif(-100, 100),
    ###
    double_h ~ dunif(-100, 100),
    double_a ~ dunif(-100, 100),
    ###
    triple_h ~ dunif(-100, 100),
    triple_a ~ dunif(-100, 100),
    ###
    HR_h ~ dunif(-100, 100),
    HR_a ~ dunif(-100, 100),
    ###
    balls_h ~ dunif(-100, 100),
    balls_a ~ dunif(-100, 100),
    ###
    hits_allowed_h ~ dunif(-100, 100),
    hits_allowed_a ~ dunif(-100, 100),
    ###
    pballs_h ~ dunif(-100, 100),
    pballs_a ~ dunif(-100, 100),
    ###
    pstrikeouts_h ~ dunif(-100, 100),
    pstrikeouts_a ~ dunif(-100, 100),
    ###
    strikes_h ~ dunif(-100, 100),
    strikes_a ~ dunif(-100, 100),
    ###standard priors
    c(fh1, fh2, fh3, fh4, fa1, fa2, fa3, fa4) ~ dunif(-100, 100),
    sigma ~ dcauchy(0,2)),
  data=TEMP,
  iter=100, warmup=10, chains=1, cores=4, control = list(adapt_delta = 0.99, max_treedepth = 15))
```

```{r varint.m, eval=FALSE, echo=FALSE}
set.seed(1234)
options(mc.cores = parallel::detectCores())
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')
###
varint.m <- map2stan(
  alist(
    rdiff ~ dnorm( mu, sigma),
    mu <- A + B + C + D,
    A <- a + h[home] + a[away],
    B <- b + hits_h * Hh + double_h * B2Bh + triple_h * B3Bh + HR_h * BHRh + balls_h * BBBh +
             hits_a * Ha + double_a * B2Ba + triple_a * B3Ba + HR_a * BHRa + balls_a * BBBa,
    C <- c + hits_allowed_h * HHAh + pballs_h * PBBh + pstrikeouts_h * PSOh +  strikes_h * Strh +
             hits_allowed_a * HHAa + pballs_a * PBBa + pstrikeouts_a * PSOa + strikes_a * Stra,
    D <- d + fh1 * PO_h + fh2 * A_h + fh3 * E_h + fh4 * DP_h + fa1 * PO_a + fa2 * A_a + fa3 * E_a + fa4 * DP_a,
    ###adaptive priors for coefficients
    h[home] ~ dnorm(h_mu, h_sigma),
    h_mu ~ dnorm(0,1),
    h_sigma ~  dcauchy(0,2), 
    a[away] ~ dnorm(a_mu, a_sigma),
    a_mu ~ dnorm(0,1),
    a_sigma ~ dcauchy(0,2),
    ###
    hits_h ~ dnorm(hits_h_mu, hits_h_sigma),
    hits_h_mu ~ dnorm(0,1),
    hits_h_sigma ~ dcauchy(0,2), 
    hits_a ~ dnorm(hits_a_mu, hits_a_sigma),
    hits_a_mu ~ dnorm(0,1),
    hits_a_sigma ~ dcauchy(0,2),
    ###
    double_h ~ dnorm(double_h_mu, double_h_sigma),
    double_h_mu ~ dnorm(0,1),
    double_h_sigma ~ dcauchy(0,2),
    double_a ~ dnorm(double_a_mu, double_a_sigma),
    double_a_mu ~ dnorm(0,1),
    double_a_sigma ~ dcauchy(0,2),    
    ###
    triple_h ~ dnorm(triple_h_mu, triple_h_sigma),
    triple_h_mu ~ dnorm(0,1),
    triple_h_sigma ~ dcauchy(0,2),
    triple_a ~ dnorm(triple_a_mu, triple_a_sigma),
    triple_a_mu ~ dnorm(0,1),
    triple_a_sigma ~ dcauchy(0,2),
    ###
    HR_h ~ dnorm(HR_h_mu, HR_h_sigma),
    HR_h_mu ~ dnorm(0,1),
    HR_h_sigma ~ dcauchy(0,2),
    HR_a ~ dnorm(HR_a_mu, HR_a_sigma),
    HR_a_mu ~ dnorm(0,1),
    HR_a_sigma ~ dcauchy(0,2),    
    ###
    balls_h ~ dnorm(balls_h_mu, balls_h_sigma),
    balls_h_mu ~ dnorm(0,1),
    balls_h_sigma ~ dcauchy(0,2),
    balls_a ~ dnorm(balls_a_mu, balls_a_sigma),
    balls_a_mu ~ dnorm(0,1),
    balls_a_sigma ~ dcauchy(0,2),      
    ###
    hits_allowed_h ~ dnorm(hits_allowed_h_mu, hits_allowed_h_sigma),
    hits_allowed_h_mu ~ dnorm(0,1),
    hits_allowed_h_sigma ~ dcauchy(0,2),
    hits_allowed_a ~ dnorm(hits_allowed_a_mu, hits_allowed_a_sigma),
    hits_allowed_a_mu ~ dnorm(0,1),
    hits_allowed_a_sigma ~ dcauchy(0,2),      
    ###
    pballs_h ~ dnorm(pballs_h_mu, pballs_h_sigma),
    pballs_h_mu ~ dnorm(0,1),
    pballs_h_sigma ~ dcauchy(0,2),
    pballs_a ~ dnorm(pballs_a_mu, pballs_a_sigma),
    pballs_a_mu ~ dnorm(0,1),
    pballs_a_sigma ~ dcauchy(0,2),      
    ###
    pstrikeouts_h ~ dnorm(pstrikeouts_h_mu, pstrikeouts_h_sigma),
    pstrikeouts_h_mu ~ dnorm(0,1),
    pstrikeouts_h_sigma ~ dcauchy(0,2),
    pstrikeouts_a ~ dnorm(pstrikeouts_a_mu, pstrikeouts_a_sigma),
    pstrikeouts_a_mu ~ dnorm(0,1),
    pstrikeouts_a_sigma ~ dcauchy(0,2),     
    ###
    strikes_h ~ dnorm(strikes_h_mu, strikes_h_sigma),
    strikes_h_mu ~ dnorm(0,1),
    strikes_h_sigma ~ dcauchy(0,2),
    strikes_a ~ dnorm(strikes_a_mu, strikes_a_sigma),
    strikes_a_mu ~ dnorm(0,1),
    strikes_a_sigma ~ dcauchy(0,2),
    ###standard priors
    c(a, b, c, d) ~ dnorm(0,1),
    ###  
    c(fh1, fh2, fh3, fh4, fa1, fa2, fa3, fa4) ~ dnorm(0,1),
    sigma ~ dcauchy(0,2)),
  data=mydata,
  iter=5000, warmup=500, chains=1, cores=4,
  control = list(adapt_delta = 0.99, max_treedepth = 15))

```

nonmlm.m and varint.m object are not in environment. 

```{r modelcompareplot, eval=FALSE}
obj <- compare(nonmlm.m, varint.m, vivs.m)
plot(compare(nonmlm.m, varint.m, vivs.m))
compare(nonmlm.m, varint.m, vivs.m)

str(obj)
TEMP <- NULL
TEMP$WAIC <-obj$WAIC 
TEMP$SE <- obj$SE
TEMP$pWAIC <- obj$WAIC - obj$pWAIC 

TEMP <- as.data.frame(TEMP)
TEMP$MODEL <- c("Varying Intercept Only", "Varying Intercept & Slope", "Standard Regression")

TEMP <- TEMP[1:2,]
TEMP
colnames(TEMP) <- c("Out of Sample", "SE", "In Sample", "MODEL")
mdata <- melt(TEMP, id=c("MODEL", 'SE'))
mdata
colnames(mdata) <- c("MODEL", "SE", "Deviance", "value")
'ggplot(TEMP, aes(y=MODEL)) +
  geom_point(shape = 1, aes(x=WAIC)) +
  geom_segment((aes(xend=WAIC, yend=MODEL, x=2*SE + WAIC))) + 
  geom_segment((aes(xend=WAIC, yend=MODEL, x=WAIC-2*SE))) +
  geom_point(aes(x=pWAIC)) +
  theme_minimal() +
  labs(title="Comparing Model Deviance") +
    xlab("Deviance") + ylab("") +
  theme(legend.position="bottom") '

mdata
mdata$MODEL <- as.factor(mdata$MODEL)
#levels(mdata$MODEL) <- gsub(" ", "\n", levels(mdata$MODEL))
levels(mdata$MODEL) <- c("Varying\nIntercept\nOnly", "Varying\nIntercept\n& Slope")
mdata$MODEL <- relevel(mdata$MODEL, "Varying\nIntercept\n& Slope")

ggplot(mdata, aes(y=MODEL)) +
  geom_point(aes(x=value, shape = Deviance)) +
  geom_segment((aes(xend=value, yend=MODEL, x=2*SE + value)), data = mdata[mdata$Deviance=="Out of Sample", ] ) + 
  geom_segment((aes(xend=value, yend=MODEL, x=-2*SE + value)), data = mdata[mdata$Deviance=="Out of Sample", ] ) + 
  theme_minimal() +
  labs(title="Comparing Model Deviance") +
    xlab("Deviance") + ylab("") +
  scale_shape_manual(values = c(1, 19)) +
  theme(legend.position="bottom") 

```

###Posterior Distributions  

Simulate game 1 with team and fielding effects only.

```{r game1data, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
wspre <- read_csv("data/WS Investigation/wsprezero.csv")
wspre <- as.data.frame(wspre)
wspre$home <- as.factor(wspre$home)
wspre$away <- as.factor(wspre$away)
sim.model1 <- sim(vivs.m, data=as.data.frame(wspre), n=750)
```

Simulate game 2 with game 1 data and new model.

```{r game2modeldata, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
ws <- read_csv("data/WS Investigation/ws.csv")
ws <- as.data.frame(ws)
namesws <- colnames(ws)
attributes(ws) <- NULL
ws <- as.data.frame(ws)
colnames(ws) <- namesws
colnames(ws)[1] <- "X"
###
TEMP <- ws[1,] #rdiff is how they performed as the home team
TEMP2 <- rbind(mydata, TEMP)
```
```{r game2model, eval=FALSE, echo=FALSE}
set.seed(1234)
options(mc.cores = parallel::detectCores())
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')

game2.m <- map2stan(
  alist(
    rdiff ~ dnorm( mu, sigma),
    ###likelihood
    mu <- A + BH + PH + BA + PA + FH + FA,
    #varying intercepts
    A <- h[home] + a[away] + a, 
    #home- batting
    BH <- hits_h[home] * Hh + double_h[home] * B2Bh + triple_h[home] * B3Bh + 
         HR_h[home] * BHRh + balls_h[home] * BBBh + bh, 
    #home- pitching
    PH <-  hits_allowed_h[home] * HHAh + pballs_h[home] * PBBh + 
          pstrikeouts_h[home] * PSOh +  strikes_h[home] * Strh + ph,
    #away- batting
    BA <- hits_a[away] * Ha + double_a[away] * B2Ba + triple_a[away] * B3Ba + 
      HR_a[away] * BHRa + balls_a[away] * BBBa +  ba, 
    #away- pitching
    PA <- hits_allowed_a[away] * HHAa + pballs_a[away] * PBBa + 
      pstrikeouts_a[away] * PSOa + strikes_a[away] * Stra + pa,
    #home- fielding
    FH <- fh1 * PO_h + fh2 * A_h + fh3 * E_h + fh4 * DP_h + fh,
    FA <-  fa1 * PO_a + fa2 * A_a + fa3 * E_a + fa4 * DP_a + fa,
    ###adaptive priors
    c(h, hits_h, double_h, triple_h, HR_h, balls_h, hits_allowed_h, pballs_h, 
      pstrikeouts_h, strikes_h)[home] ~ dmvnormNC(sigma_home,Rho_home),
    c(a, hits_a, double_a, triple_a, HR_a, balls_a, hits_allowed_a, pballs_a, 
      pstrikeouts_a, strikes_a)[away] ~ dmvnormNC(sigma_away,Rho_away),
    ###standard priors
    c(a, bh, ph, fh, ba, pa, fa) ~ dnorm(0,1), #hyperpriors
    c(fh1, fh2, fh3, fh4, fa1, fa2, fa3, fa4) ~ dnorm(0,1), #scaled variables
    sigma_home ~ dcauchy(0,2),
    sigma_away ~ dcauchy(0,2),
    Rho_home ~ dlkjcorr(4),
    Rho_away ~ dlkjcorr(4),
    sigma ~ dcauchy(0,2)),
  data=TEMP2,
  iter=1000, warmup=250, chains=1, cores=4, control = list(adapt_delta = 0.99, max_treedepth = 15))
```
```{r game2prob, warning=FALSE, echo=FALSE, include=FALSE}
sim.model2 <- sim(game2.m, data=as.data.frame(TEMP), n=750)
```

Simulate game 3 with LAD now as home team- like game 1, use team and fielding effects.

```{r game3modeldata, echo=FALSE}
TEMP <- ws[1:2,] #rdiff is how they performed as the home team
TEMP2 <- rbind(mydata, TEMP)
```
```{r game3model, eval=FALSE, echo=FALSE}
set.seed(1234)
options(mc.cores = parallel::detectCores())
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')

game3.m <- map2stan(
  alist(
    rdiff ~ dnorm( mu, sigma),
    ###likelihood
    mu <- A + BH + PH + BA + PA + FH + FA,
    #varying intercepts
    A <- h[home] + a[away] + a, 
    #home- batting
    BH <- hits_h[home] * Hh + double_h[home] * B2Bh + triple_h[home] * B3Bh + 
         HR_h[home] * BHRh + balls_h[home] * BBBh + bh, 
    #home- pitching
    PH <-  hits_allowed_h[home] * HHAh + pballs_h[home] * PBBh + 
          pstrikeouts_h[home] * PSOh +  strikes_h[home] * Strh + ph,
    #away- batting
    BA <- hits_a[away] * Ha + double_a[away] * B2Ba + triple_a[away] * B3Ba + 
      HR_a[away] * BHRa + balls_a[away] * BBBa +  ba, 
    #away- pitching
    PA <- hits_allowed_a[away] * HHAa + pballs_a[away] * PBBa + 
      pstrikeouts_a[away] * PSOa + strikes_a[away] * Stra + pa,
    #home- fielding
    FH <- fh1 * PO_h + fh2 * A_h + fh3 * E_h + fh4 * DP_h + fh,
    FA <-  fa1 * PO_a + fa2 * A_a + fa3 * E_a + fa4 * DP_a + fa,
    ###adaptive priors
    c(h, hits_h, double_h, triple_h, HR_h, balls_h, hits_allowed_h, pballs_h, 
      pstrikeouts_h, strikes_h)[home] ~ dmvnormNC(sigma_home,Rho_home),
    c(a, hits_a, double_a, triple_a, HR_a, balls_a, hits_allowed_a, pballs_a, 
      pstrikeouts_a, strikes_a)[away] ~ dmvnormNC(sigma_away,Rho_away),
    ###standard priors
    c(a, bh, ph, fh, ba, pa, fa) ~ dnorm(0,1), #hyperpriors
    c(fh1, fh2, fh3, fh4, fa1, fa2, fa3, fa4) ~ dnorm(0,1), #scaled variables
    sigma_home ~ dcauchy(0,2),
    sigma_away ~ dcauchy(0,2),
    Rho_home ~ dlkjcorr(4),
    Rho_away ~ dlkjcorr(4),
    sigma ~ dcauchy(0,2)),
  data=TEMP2,
  iter=1000, warmup=250, chains=1, cores=4, control = list(adapt_delta = 0.99, max_treedepth = 15))
```
```{r game3simdata, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
wsg3pre <- read_csv("data/WS Investigation/wsg3pre.csv")
wsg3pre <- as.data.frame(wsg3pre)
nameswsg3pre <- colnames(wsg3pre)
attributes(wsg3pre) <- NULL
wsg3pre <- as.data.frame(wsg3pre)
colnames(wsg3pre) <- nameswsg3pre
colnames(wsg3pre)[1] <- "X"
```
```{r game3prob, echo=FALSE, include=FALSE}
sim.model3 <- sim(game3.m, data=as.data.frame(wsg3pre), n=750)
```

Simulate game 4 with game 3 data and new model.

```{r game4modeldata, echo=FALSE, include=FALSE}
TEMP <- ws[1:3,] #rdiff is how they performed as the home team
TEMP2 <- rbind(mydata, TEMP)
```
```{r game4model, eval=FALSE, echo=FALSE}
set.seed(1234)
options(mc.cores = parallel::detectCores())
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')

game4.m <- map2stan(
  alist(
    rdiff ~ dnorm( mu, sigma),
    ###likelihood
    mu <- A + BH + PH + BA + PA + FH + FA,
    #varying intercepts
    A <- h[home] + a[away] + a, 
    #home- batting
    BH <- hits_h[home] * Hh + double_h[home] * B2Bh + triple_h[home] * B3Bh + 
         HR_h[home] * BHRh + balls_h[home] * BBBh + bh, 
    #home- pitching
    PH <-  hits_allowed_h[home] * HHAh + pballs_h[home] * PBBh + 
          pstrikeouts_h[home] * PSOh +  strikes_h[home] * Strh + ph,
    #away- batting
    BA <- hits_a[away] * Ha + double_a[away] * B2Ba + triple_a[away] * B3Ba + 
      HR_a[away] * BHRa + balls_a[away] * BBBa +  ba, 
    #away- pitching
    PA <- hits_allowed_a[away] * HHAa + pballs_a[away] * PBBa + 
      pstrikeouts_a[away] * PSOa + strikes_a[away] * Stra + pa,
    #home- fielding
    FH <- fh1 * PO_h + fh2 * A_h + fh3 * E_h + fh4 * DP_h + fh,
    FA <-  fa1 * PO_a + fa2 * A_a + fa3 * E_a + fa4 * DP_a + fa,
    ###adaptive priors
    c(h, hits_h, double_h, triple_h, HR_h, balls_h, hits_allowed_h, pballs_h, 
      pstrikeouts_h, strikes_h)[home] ~ dmvnormNC(sigma_home,Rho_home),
    c(a, hits_a, double_a, triple_a, HR_a, balls_a, hits_allowed_a, pballs_a, 
      pstrikeouts_a, strikes_a)[away] ~ dmvnormNC(sigma_away,Rho_away),
    ###standard priors
    c(a, bh, ph, fh, ba, pa, fa) ~ dnorm(0,1), #hyperpriors
    c(fh1, fh2, fh3, fh4, fa1, fa2, fa3, fa4) ~ dnorm(0,1), #scaled variables
    sigma_home ~ dcauchy(0,2),
    sigma_away ~ dcauchy(0,2),
    Rho_home ~ dlkjcorr(4),
    Rho_away ~ dlkjcorr(4),
    sigma ~ dcauchy(0,2)),
  data=TEMP2,
  iter=1000, warmup=250, chains=1, cores=4, control = list(adapt_delta = 0.99, max_treedepth = 15))
```
```{r game4prob, echo=FALSE, include=FALSE}
TEMP <- ws[3,] 
sim.model4 <- sim(game4.m, data=as.data.frame(TEMP), n=750)
```

Simulate game 5 with games 3 and 4 data and new model.
```{r game5modeldata, echo=FALSE, include=FALSE}
TEMP <- ws[1:4,] #rdiff is how they performed as the home team
TEMP2 <- rbind(mydata, TEMP)
```
```{r game5model, eval=FALSE, echo=FALSE, message=FALSE}
set.seed(1234)
options(mc.cores = parallel::detectCores())
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')

game5.m <- map2stan(
  alist(
    rdiff ~ dnorm( mu, sigma),
    ###likelihood
    mu <- A + BH + PH + BA + PA + FH + FA,
    #varying intercepts
    A <- h[home] + a[away] + a, 
    #home- batting
    BH <- hits_h[home] * Hh + double_h[home] * B2Bh + triple_h[home] * B3Bh + 
         HR_h[home] * BHRh + balls_h[home] * BBBh + bh, 
    #home- pitching
    PH <-  hits_allowed_h[home] * HHAh + pballs_h[home] * PBBh + 
          pstrikeouts_h[home] * PSOh +  strikes_h[home] * Strh + ph,
    #away- batting
    BA <- hits_a[away] * Ha + double_a[away] * B2Ba + triple_a[away] * B3Ba + 
      HR_a[away] * BHRa + balls_a[away] * BBBa +  ba, 
    #away- pitching
    PA <- hits_allowed_a[away] * HHAa + pballs_a[away] * PBBa + 
      pstrikeouts_a[away] * PSOa + strikes_a[away] * Stra + pa,
    #home- fielding
    FH <- fh1 * PO_h + fh2 * A_h + fh3 * E_h + fh4 * DP_h + fh,
    FA <-  fa1 * PO_a + fa2 * A_a + fa3 * E_a + fa4 * DP_a + fa,
    ###adaptive priors
    c(h, hits_h, double_h, triple_h, HR_h, balls_h, hits_allowed_h, pballs_h, 
      pstrikeouts_h, strikes_h)[home] ~ dmvnormNC(sigma_home,Rho_home),
    c(a, hits_a, double_a, triple_a, HR_a, balls_a, hits_allowed_a, pballs_a, 
      pstrikeouts_a, strikes_a)[away] ~ dmvnormNC(sigma_away,Rho_away),
    ###standard priors
    c(a, bh, ph, fh, ba, pa, fa) ~ dnorm(0,1), #hyperpriors
    c(fh1, fh2, fh3, fh4, fa1, fa2, fa3, fa4) ~ dnorm(0,1), #scaled variables
    sigma_home ~ dcauchy(0,2),
    sigma_away ~ dcauchy(0,2),
    Rho_home ~ dlkjcorr(4),
    Rho_away ~ dlkjcorr(4),
    sigma ~ dcauchy(0,2)),
  data=TEMP2,
  iter=1000, warmup=250, chains=1, cores=4, control = list(adapt_delta = 0.99, max_treedepth = 15))
```
```{r game5prob, echo=FALSE, include=FALSE}
TEMP <- ws[3:4,] 
sim.model5 <- sim(game5.m, data=as.data.frame(TEMP), n=750)
```

```{r predictiondataset, echo=FALSE}
TEMP <- NULL
TEMP1 <- -1 * as.data.frame(sim.model1) #predicting game 1- make negative for probability LAD (away) win greater than 0

TEMP2 <- -1* as.data.frame(sim.model2) #predicting game 2- make negative for probability LAD (away) win greater than 0
colnames(TEMP2) <- c("V2")

TEMP3 <- as.data.frame(sim.model3) #predicting game 3
TEMP3$V1 = TEMP3$V1
colnames(TEMP3) <- c("V3")

TEMP4 <- as.data.frame(sim.model4) #predicting game 4 (LA won game 3)
TEMP4$V1 = TEMP4$V1
colnames(TEMP4) <- c("V4")

TEMP5 <- as.data.frame(sim.model5) #predicting game 5
TEMP5$V1 = TEMP5$V1
colnames(TEMP5) <- c("V5")

TEMP <- cbind(TEMP1, TEMP2, TEMP3, TEMP4, TEMP5)
```

```{r posteriorplot, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(TEMP) +
    geom_freqpoly(aes(x=V1, y=..density.., alpha=0.6), col='#c6dbef', size=1.2) +
    geom_freqpoly(aes(x=V2, y=..density.., alpha=0.7), col='#9ecae1', size=1.2) + #following BOS first win
    geom_freqpoly(aes(x=V3, y=..density.., alpha=0.8), col='#6baed6', size=1.2) +
    geom_freqpoly(aes(x=V4, y=..density.., alpha=0.9), col='#3182bd', size=1.2) +
    geom_freqpoly(aes(x=V5, y=..density.., alpha=1), col='#08519c', size=1.2) +
    theme_minimal() + guides(fill=FALSE) + 
    labs(title="World Series 2018: Posterior Predictions Games 1-5", subtitle="Positive x-values indicate LAD Win") +
    xlab("Difference in Points (LAD - BOS)") +
  scale_alpha_continuous(labels = c("Game 1", "Game 2", "Game 3", "Game 4", "Game 5")) +
  theme(legend.position="bottom") +
  guides(alpha=guide_legend("")) +
  theme(plot.title = element_text(size=18), plot.subtitle = element_text(size=14),
        axis.text.x = element_text(size=12), axis.title.x = element_text(size=12),
        legend.text=element_text(size=12))

```

###Pooled unPooled

```{r CILB, echo=FALSE}
pred <- NULL
pred <- data.frame("LB" = c(
  quantile(TEMP1$V1, 0.025), #game 1 prediction- slightly favor BOS
  quantile(TEMP2$V2, 0.025), #game 2 prediction- strong favor BOS
  quantile(TEMP3$V3, 0.025), #game 3 prediction- moderate favor BOS
  quantile(TEMP4$V4, 0.025), #game4 prediction (LA won game 3)- stronger favor BOS
  quantile(TEMP5$V5, 0.025)) #game 5 prediction- mild favor BOS
  ,
  "UB" = c(
  quantile(TEMP1$V1, 0.975), #game 1 prediction- slightly favor BOS
  quantile(TEMP2$V2, 0.975), #game 2 prediction- strong favor BOS
  quantile(TEMP3$V3, 0.975), #game 3 prediction- moderate favor BOS
  quantile(TEMP4$V4, 0.975), #game4 prediction (LA won game 3)- stronger favor BOS
  quantile(TEMP5$V5, 0.975))
  ,
  "Mean" = c(
  mean(TEMP1$V1), #game 1 prediction- slightly favor BOS
  mean(TEMP2$V2), #game 2 prediction- strong favor BOS
  mean(TEMP3$V3), #game 3 prediction- moderate favor BOS
  mean(TEMP4$V4), #game4 prediction (LA won game 3)- stronger favor BOS
  mean(TEMP5$V5)) #game 5 prediction- mild favor BOS  
  ,
  "Game" = c(1:5)) 
```

```{r}
hometeam = "LAD"
TEMP <- mydata %>% filter(home == hometeam)
mean(TEMP$rdiff)

awayteam = "LAD"
TEMP <- mydata %>% filter(away == awayteam)
-1 * mean(TEMP$rdiff)

pred$pool <- c(1.56, 1.56, .84, .84, .84)
```

```{r}
pred$nonmlm <- c()
```


```{r pooledunpooled}
par(mfrow=c(1,1))
plot(pred$Mean, pred$Game, type = "n", xlim=c(-10, 10), main ="95% CI, Mean, Pooling")
segments(pred$Mean + pred$LB, pred$Game, pred$Mean + pred$UB, pred$Game)
points(pred$Mean, pred$Game, col = "red", pch=4) 
points(pred$pool, pred$Game, col = "green", pch=4) 

head(pred)

pred$unpooled <- c(.84, .84, 2.3, 2.3, 2.3)
pred$actual <- c(-4, -2, 1, -3, -4)

head(pred)

pred2 <- NULL
pred2$Mean = c(-0.444, -3.796, 1.26, 1, 1.2, 0.84, 0.84, 2.3, 2.3, 2.3, -4, -5, .5, -3, -4)
pred2$Game = c(1, 2, 3, 4, 5, 1, 2, 3, 4, 5, 1, 2, 3, 4, 5)
pred2$LB = rep(pred$LB, 3)
pred2$UB = rep(pred$UB, 3)
pred2$Model <- c(rep(c("Partial Pooling"), 5), rep(c("Pooled"), 5), rep(c("Unpooled"), 5))
pred2 <- as.data.frame(pred2)
head(pred2)

ggplot(pred2, aes(y=as.factor(Game), pch=Model)) +
  geom_point(aes(x=Mean), size=3) +
  scale_shape_manual(values=c(4, 1, 3))+
  geom_segment((aes(xend=Mean, yend=Game, x=1.96*LB, col=Game, alpha=Game)), size=1) +
  geom_segment((aes(xend=Mean, yend=Game, x=1.96*UB, col = Game, alpha=Game)), size=1) + 
  geom_point(aes(x=Mean), size=3) +
  scale_shape_manual(values=c(4, 1, 3))+
  theme_minimal() +
  labs(title="Partial Pooling with Multi-level Model") +
  xlab("Difference in Points (LAD - BOS)") + ylab("World Series Game") +
  theme(legend.position="bottom") + xlim(-20, 20) + guides(alpha=FALSE) + guides(col=FALSE) +
  theme(plot.title = element_text(size=18), plot.subtitle = element_text(size=14),
        axis.text.x = element_text(size=12), axis.title.x = element_text(size=12),
        legend.text=element_text(size=12))
```


```{r pvalues}
prob_success1 <- 1- sum(sim.model1>0) / length(sim.model1); prob_success1 #subtract from 1 for LAD away
prob_success2 <- 1- sum(sim.model2>0) / length(sim.model2); prob_success2 #subtract from 1 for LAD away
prob_success3 <- sum(sim.model3>0) / length(sim.model3); prob_success3
prob_success4 <- sum(sim.model4>0) / length(sim.model4); prob_success4
prob_success5 <- sum(sim.model5>0) / length(sim.model5); prob_success5
```

```{r binomprob, echo=FALSE, include=FALSE}
dbinom(4, size=7, prob=prob_success1) + dbinom(5, size=7, prob=prob_success1) + dbinom(6, size=7, prob=prob_success1) + dbinom(7, size=7, prob=prob_success1)
dbinom(4, size=6, prob=prob_success2) + dbinom(5, size=6, prob=prob_success2) + dbinom(6, size=6, prob=prob_success2)
dbinom(4, size=5, prob=prob_success3) + dbinom(5, size=5, prob=prob_success3) 
dbinom(3, size=4, prob=prob_success4) + dbinom(4, size=4, prob=prob_success4)
dbinom(3, size=3, prob=prob_success5)
```

###Parametric Analysis

```{r game5data2, warning=FALSE, message=FALSE, echo=FALSE, message=FALSE}
###read in pre-game data
wsg5prepbp <- read_csv("data/WS Investigation/wsg5prepbp.csv")
wspre <- wsg5prepbp
###if NA's, try below
wspre <- as.data.frame(wspre)
wspre$home <- as.factor(wspre$home)
wspre$away <- as.factor(wspre$away)
```

```{r game5parametric, echo=FALSE, message=FALSE}
###game 5 parametric
TEMP <- wspre[1,] 
link.model <- link(game5.m, data=data.frame(TEMP))
mean(link.model$BH) + mean(link.model$BA)
mean(link.model$PH) + mean(link.model$PA)
mean(link.model$FH) + mean(link.model$FA)
mean(link.model$A)
mean(link.model$mu)
.12+.14+.17

#Team Effects
#BOS (away)
BOSA <- melt(link.model$A); colnames(BOSA) <- c("Category", "x2bat", "Value")
BOSA$Category = rep("BOS", nrow(BOSA)); 
BOSA <- BOSA[,c("Category", "Value")]
BOSA$Value = BOSA$Value * -1  #for visualization
#LAD (home)
LADH <- melt(link.model$A); colnames(LADH) <- c("Category", "x2pit", "Value")
LADH$Category = rep("LAD", nrow(LADH))
LADH <- LADH[,c("Category", "Value")]
LADH$Value = LADH$Value 
#Batting Effects
#LAD home bat
LADBH <- melt(link.model$BH); colnames(LADBH) <- c("Category", "x2bat", "Value")
LADBH$Category = rep("LAD", nrow(LADBH)); 
LADBH <- LADBH[,c("Category", "Value")]
#head(LADBH); str(LADBH)
#BOS away bat
BOSBA <- melt(link.model$BA); colnames(BOSBA) <- c("Category", "x2pit", "Value")
BOSBA$Category = rep("BOS", nrow(BOSBA))
BOSBA <- BOSBA[,c("Category", "Value")]
BOSBA$Value = BOSBA$Value * -1 #for visualization
#Pitching Effects
#LAD home pit
LADPH <- melt(link.model$PH); colnames(LADPH) <- c("Category", "x2bat", "Value")
LADPH$Category = rep("LAD", nrow(LADPH)); 
LADPH <- LADPH[,c("Category", "Value")]
LADPH$Value = LADPH$Value  
#BOS away pit
BOSPA <- melt(link.model$PA); colnames(BOSPA) <- c("Category", "x2bat", "Value")
BOSPA$Category = rep("BOS", nrow(BOSPA)); 
BOSPA <- BOSPA[,c("Category", "Value")]
BOSPA$Value = BOSPA$Value * -1 #for visualization
#Fielding Effects
#BOS fielding away
BOSFA <- melt(link.model$FA); colnames(BOSFA) <- c("Category", "x2bat", "Value")
BOSFA$Category = rep("BOS", nrow(BOSFA)); 
BOSFA <- BOSFA[,c("Category", "Value")]
BOSFA$Value = BOSFA$Value * -1 #for visualization
#LAD fielding home
LADFH <- melt(link.model$FH); colnames(LADFH) <- c("Category", "x2pit", "Value")
LADFH$Category = rep("LAD", nrow(LADFH))
LADFH <- LADFH[,c("Category", "Value")]
LADFH$Value = LADFH$Value 

TEMP1 <- as.data.frame(rbind(BOSA, LADH)) #, 
TEMP1$Group <- rep("Team", nrow(TEMP1))
TEMP2 <- as.data.frame(rbind(BOSBA, LADBH)) #,
TEMP2$Group <- rep("Batting", nrow(TEMP2))
TEMP3 <- as.data.frame(rbind(BOSPA, LADPH)) #,
TEMP3$Group <- rep("Pitching", nrow(TEMP3))
TEMP4 <- as.data.frame(rbind(BOSFA, LADFH)) #,
TEMP4$Group <- rep("Fielding", nrow(TEMP4))
TEMP5 <- rbind(TEMP1, TEMP2, TEMP3, TEMP4)

ggplot(TEMP5) + 
  geom_density(aes(x = Value, group=Category, col=Category), kernel="gaussian") +
  facet_wrap(vars(Group), scales="free") +
  theme_minimal() +
  labs(title="Game 5: Parametric Analysis", 
       subtitle="Posterior Distributions of Difference in Runs For Each Parameter") +
    xlab("Difference in Points") +
  theme(legend.position="bottom") +
  scale_color_manual( values = c("#a50f15","#08519c")) +
    theme(plot.title = element_text(size=16), plot.subtitle = element_text(size=14),
        axis.text.x = element_text(size=12), axis.title.x = element_text(size=12),
        legend.text=element_text(size=12))
```

###Real-Time Modeling: Game 5

```{r, echo=FALSE, message=FALSE, include=FALSE}
surviv <- NULL
surviv[1] <- prob_success5
```
```{r firstinningfirsthalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[2,] #rdiff is how they performed as the home team
head(TEMP)
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[2] <- prob_success
```
```{r firstinning2ndhalf, echo=FALSE, message=FALSE, include=FALSE}
#BOS starts with 2 runs.
TEMP <- wspre[3,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[3] <- prob_success
```

```{r secondinningfirsthalf, echo=FALSE, message=FALSE, include=FALSE}
#LAD comes back with 1 run.
TEMP <- wspre[4,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[4] <- prob_success
```
```{r secondinningsecondhalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[5,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[5] <- prob_success
```

```{r thirdinningfirsthalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[6,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[6] <- prob_success
```
```{r thirdinningsecondhalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[7,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[7] <- prob_success
```

```{r fourthinningprior, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[8,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[8] <- prob_success
```

```{r fourthinningsecondhalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[9,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[9] <- prob_success
```

```{r fifthinningprior, echo=FALSE, message=FALSE, include=FALSE}
#BAM! Boston scores another point.
TEMP <- wspre[10,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[10] <- prob_success
```
```{r fifthinningsecondhalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[11,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[11] <- prob_success
```

```{r sixthinningpiror, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[12,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[12] <- prob_success
```
```{r sixthinningsecondhalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[13,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[13] <- prob_success
```

```{r seventhinningprior, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[14,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[14] <- prob_success
```
```{r seventhinningsecondhalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[15,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[15] <- prob_success
```

```{r eighthinningprior, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[16,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[16] <- prob_success
```
```{r eighthinningsecondhalf, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[17,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[17] <- prob_success
```

```{r ninthinningprior, echo=FALSE, message=FALSE, include=FALSE}
TEMP <- wspre[18,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[18] <- prob_success
```
```{r ninthinningsecond, echo=FALSE, message=FALSE, include=FALSE, eval=FALSE}
#Drop is BOS batting average, model doesn't assume ending at 9 innings.
TEMP <- wspre[19,] #rdiff is how they performed as the home team
sim.model <- sim(game5.m, data=as.data.frame(TEMP), n=750)
prob_success <- sum(sim.model>0) / length(sim.model) 
surviv[19] <- prob_success
```

```{r survivalgraph, echo=FALSE}
TEMP <- as.data.frame(surviv)
TEMP$inning <- seq(from=1, to=9.5, by=0.5)

ggplot(TEMP, aes(inning, surviv)) +
  geom_step(direction="hv", col='#08519c', size=1) +
  theme_minimal() +
  labs(title="Game 5: Play-by-Play Probability Updates", subtitle="Probability of LAD Game 5 Win") +
    xlab("Inning") + ylab('Probably of LAD Win') + scale_x_continuous(breaks=0:9) +
    theme(plot.title = element_text(size=16), plot.subtitle = element_text(size=14),
        axis.text.x = element_text(size=12), axis.title.x = element_text(size=12),
        legend.text=element_text(size=12))
```

